Using Visual Data Management System (VDMS) to\\ Index and Search 100M Images

### 2020-02-25 yfcc100m/paper/vldb2020/eval.tex
Update to overleaf version

We have used the YFCC100M dataset to evaluate different aspects of our system.
We use the images in the dataset and its associated metadata to implement
an image-search engine based on properties associated to those images.
This is a very common use-case we have encountered when attacking
applications such as smart-retail, sports applications, and video summarization,
where the starting point is a large set of data that must be curated
before proceeding with the data processing and neural network training.
In order to evaluate the different aspects of the performance on the
image search, we built a baseline following the methodology used in the industry, and following what we have done in the past in order to solve
the data search problem, which we describe in Section \ref{images}.

We also cover evaluation we have done using the Video interface VDMS offers,
in order to asses how the implementation performs and scale when used
for video handling.
When it comes to video, ad-hoc solutions have a large number of parameters that
can be tunned, and complex ways to handle the data. Together with that,
there is no system that enables transactional operations over videos files
in the way VDMS does. Because of the two reasons above mentioned,
we focus our efforts in understanding variations in the the performance of
VDMS and its scalability, rather than comparing it to a baseline that
would no represent a fair comparison for either of the systems.
We describe the results in Section \ref{videos}.

Similarly, we explore the behavior of the feature vector functionality in VDMS,
and an evaluation of the different trade-offs that the systems offers for
applications developers. For this, we implemented an image-search application
based on \textit{similarity} search, which we describe in Section \ref{features}

@@

The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a large
collection of 100 million public Flickr media objects created to provide free,
sharable multimedia data for research. This dataset contains approximately
99.2 million images and 0.8 million videos with metadata characterized by
25 fields such as the unique identifier, userid,
date the media was taken/uploaded, location in longitude/latitude coordinates,
device type the media was captured, URL to download the media object,
and the Creative Commons license type information.
The YFCC100M dataset also used a deep learning approach to generate autotags
which is a set of comma-separated concepts such as people, scenery, objects,
and animals and confidence scores generated from 1,570 trained
Caffe classifiers ~\cite{Thomee_2016}. Together with the autotag, there is a
probability associated to each tag to indicate the certainty level of the tag
in that image/frame.
We have also used feature vectors generated for every image and first frame
of every video \cite{features} to implement similarity search.
Given that there is no standard benchmark oriented towards visual data queries,
we have built a series of queries to filter this dataset that is model after
our internal use cases for many of the mentioned applications we have worked
with.

\subsection{Experimental Setup}

As a baseline, and given that there are no other open-source system
that implements similar functionality and interfaces as VDMS,
we implemented an equivalent visual data management system comprised of a
combination of widely available, off-the-shelve components:
MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
We have implemented a set of client-side applications that takes care
of retrieving the components from the different systems, and applies
pre-processing operations when needed.
For visual data workloads, building an approach like the one
implemented as a baseline for this work
is the common practice in the industry \cite{heystack, tao}.

For all our experiments, we use 2 servers, one hosting a VDMS server and
another hosting a the Baseline implementation.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with multithreading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
The client application (running the queries and measuring round-trip time)
is connected to the server through a 1GB wired link through
a 10GB-backplane switch.

Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (which wants to retrieve metadata and
imaged) with VDMS (left) and the baseline (right).
The client application was implemented using Python 3 for both VDMS and
the baseline.

It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
and VDMS have direct access. In the case of the baseline, metadata is
stored in MySQL using an attached SSD disk.
Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of comparison w.r.t
MySQL (which was not design for Persistent Memory type of storage).
The benefits of Persistent Memory on metadata operations is left
for another paper, and outside the scope of this evaluation.
For this experiment then, in the case of VDMS we simply use a similar
attached SSD disk to store metadata.
Even if PMGD, the graph database used by VDMS, is designed for PM,
it can deliver good performance when using SSDs directly.

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/comparison_system}
\caption{Comparison Systems}
\label{fig:systems}
\end{figure*}

For the metadata, we built VDMS and MySQL databases
using the YFCC100M dataset with incremental database sizes.
For simplicity, we named the database based on the approximate number of images
it contains, as follows: 100k, 500k, 1M, 5M, 10M, 50M, 100M.
The exact number of images/elements on each database is shown in
Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}
For each database size, we created an instance of VDMS using YFCC100m metadata,
the YFCC100m machine-generated \textit{autotags} associated with
each metadata identifier, and the list of 1,570 \textit{autotags}
using 100 concurrent VDMS Python clients.
Internally, that information is represented as a property graph,
where we have one node for each image, one node for each tag
(always 1570 tags), and connections between images.
For instance, if an image has 4 \textit{autotags} assigned,
there will be 4 connections between that image and
the different nodes for those \textit{autotags}.
The information about the probability the \textit{autotag} present in an image
is expressed as a property in the \textit{connection} between the 2 nodes.
Figure \ref{fig:graph_representation} shows an example on two images and
2 \textit{autotags}, and the \textit{connections} between those \textit{autotags}
and the images.
Image id: 23143252 has two \textit{autotags} assigned:
\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
Image id: 86756231, on the other hand, has a single \textit{autotags} assigned:
\textit{Alligator} with probability 0.894.

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/graph_representation}
\caption{VDMS Data Representation Using a Property Graph:
Example on two images and 2 \textit{autotags} with
their respective probabilities expressed in the \textit{connection}.
Image id: 23143252 has two \textit{autotags} assigned:
\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
Similarly, Image id: 86756231 has a single \textit{autotags} assigned:
\textit{Alligator} with probability 0.894.}
\label{fig:graph_representation}
\end{figure}

Because there are on average 8 tags assigned to each image,
there will be around 8 times more connections than images, as can be seen
in Table~\ref{table:vdmsnodes}.
Also, each image node will contain multiple properties associated
with that object.
Using the Python Client Module, we insert 1,570 nodes for the \textit{autotags},
and then insert nodes for the YFCC media objects (images or videos)
with the associated metadata.
Adding the autotag confidence scores is a recursive process and requires more
complex queries than adding the previous nodes.
For each metadata identifier in \textit{autotags},
the query must find the associated metadata node and tag node.
A connection between the two nodes is created with the confidence score as a
property. The number of metadata nodes are dependent on the database size and
the connections are responsible for 90\% of the elements in each database as
shown in Table~\ref{table:vdmsnodes}.
It is important to note that the metadata identifier, \textit{autotags}, and
longitude/latitude coordinates are set as indexes of the database to allow
faster retrieval of the metadata.

\begin{table}[h]
\caption{VDMS Database - Number of Elements}
\centering
\begin{tabular}{c c c c}
\hline\hline
DB Name & \# Images & \# Connections & \# Autotags\\
\hline
100k & 100,000     & 848,432      & 1,570\\
500k & 500,000     & 4,249,500    & 1,570\\
1M   & 1,000,000   & 8,503,045    & 1,570\\
5M   & 5,000,000   & 42,505,478   & 1,570\\
10M  & 10,000,000  & 85,040,404   & 1,570\\
50M  & 50,000,000  & 425,162,070  & 1,570\\
100M & 99,205,984  & 895,572,430  & 1,570\\
\hline
\end{tabular}
\label{table:vdmsnodes}
\end{table}

Each MySQL database is created in a similar manner but the data is represented
as three tables, following the relational model:
1) \textit{images} table, contains one row per image,
and a column for each property
associated with the images (some of which are listed in \ref{dataset});
2) \textit{taglist} table: contains one row per each autotag elements
(always 1570 rows);
3) \textit{autotags} table: contains on row per each autotag
assigned to an image, each row containing a foreign key to the
image and a foreign key to the tag, and
the probability assigned to that tag belonging to that image.
Given that there are 8 autotags, on average, per image, the \textit{autotags}
table has around 8 times the number of rows present in the
\textit{metadata} table, as can be seen in~\ref{table:mysqltables}.

THIS NEEDS FIX, NOT SURE I UNDERSTAND WHAT YOU MEAN:
By default, MySQL uses infinite number of threads inside InnoDB when processing
requests using four threads per IO read/write.
In some cases, when creating large databases data locks may occur to protect the
data from concurrent updates.
To minimize locking in our runs, we increased the InnoDB buffer pool size to
increase the amount of memory allocated to in-memory data structures
~\cite{mysql,mysql_blog}.

Using a Python client and simple queries, the \textit{taglist}
table is read from the list of tags with an auto-incremented
\textit{tagid} as a primary key and the metadata table
is read from the YFCC100m metadata using the identifier as a primary key.
The \textit{autotags} table contains the generated autotags and
probabilities for entries of the \textit{images} table.

To generate the table, we split the \textit{autotags} data for each database
by the image identifier and autotag into new files.
The new files are read into the \textit{autotags} table with the image
identifier and \textit{tagid} as foreign keys.

\begin{table}[h]
\caption{MySQL Database - Number of Rows in each Table}
\centering
\begin{tabular}{c c c c}
\hline\hline
 & \multicolumn{3}{c}{Table}\\
\cline{2-4}
DB Name & images & autotags & taglist\\
\hline
100k & 100,000    & 848,912     & 1,570\\
500k & 498,707    & 4,241,200   & 1,570\\
1M   & 1,000,000  & 8,508,380   & 1,570\\
5M   & 4,987,379  & 42,425,905  & 1,570\\
10M  & 10,000,000 & 85,095,265  & 1,570\\
50M  & 50,000,000 & 425,446,208 & 1,570\\
100M & 99,206,564 & 896,002,496 & 1,570\\
\hline
\end{tabular}
\label{table:mysqltables}
\end{table}

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/db_time_size}
\caption{Time (in hours) to build and size (in GB) of MySQL and VDMS databases}
\label{fig:db_time_size}
\end{figure}

The VDMS and MySQL databases have comparable number of elements as shown in
Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}, and the differences
can be attributed to failures in data preparation/loading because of
incomplete/inconsistent formating, which is common on large datasets.
The difference is less than 0.1\% in terms of number of elements (images and/or
metadata information).

@@

One of the first things we noticed is the difference is loading times,
where VDMS outperforms MySQL by a large margin.
Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
faster than MySQL, and how the speedup is sustained as the database size grows.
Key difference in the build times are attributed to the low-level
implementation of how MySQL reads and stores data from the files and the
optimizations (increased InnoDB pool size, etc.) needed
to handle large datasets such as YFCC100m.
On average, it took MySQL 3.72x longer to build each database than VDMS.
For example, to build the 100M database, MySQL took 263 hours
while VDMS only needed 72.5 hours.

@@

Another aspect important to note is that
VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
This is space used to store information about each node/connection.
The Graph Database internal to VDMS (called PMGD) was designed for performance,
specially in environments where persistent memory is present.
This design decision comes as a trade-off for storage footprint, which is
noticeable in our results.
VDMS required 30-41\% more storage than MySQL for storing the same amount
of metadata.
This may become a factor if storage is a limitation, but it should also be noted
that even if we have a 41\% increase in metadata size,
metadata accounts for less than 2\% of the overall database size.
Take, for example, the largest database (both metadata and images) we built
(100M): around 230GB of metadata and 12TB of images.
On the other hand, in systems were persistent memory is a scarce resource,
the increase storage foot print of PMGD may represent a challenge.

@@

\subsection{Images + Metadata}
\label{images}

In order to evaluate VDMS and the baseline on our use-case queries,
we implemented 4 queries that that filter and retrieve certain specific images.
We choose these queries because they represent typical use-case where a
cohort of images is to be retrieve and processed from a large corpus of data.
As we mentioned before, we took this approach due to the lack of a standard
benchmarks that are oriented towards visual data retrieval.
We use the metadata associated to the images to filter said images.
In particular, this dataset contains a set of 1570 machine generated
tags (\textit{autotags})assigned to the images depending on the content.
\textit{autotags} includes things like objects (chair, person, people, etc), and
concepts like places (indoor, outdoor, lake) and others (party, etc).
Together with each \textit{autotag}, a probability is assign.
This is, an image can have the \textit{autotags} people, person, party,
outdoor, and each \textit{autotag} assigned will be accompanied by a
probability of that \textit{autotags} being present in that image.
As mentioned earlier, there are 8 tags assigned to each image on average.
Also, some images have geo-location information (latitude/longitude).

We use the \textit{autotags} (as they contain information about
the content of the image), as well as the geo-location
(as an example of properties of the images used used for search and filtering).

On top of that and for our use case, we would like to extract more information
about the content of the image through the use of ML,
such as Convolutional Neural Networks TODO-CITE HERE.
For this, we resize the images to 224x224, which is the input layer size for
a popular neural network for object detection on images: ResNet TODO-CITEHERE.

To evaluate the access to metadata and images,
we use the following four queries, modeled after our internal use cases:
\begin{itemize}
\item Query 1 - {\bf {\em 1tag}}: Find metadata/images with 1 specific autotag (i.e. alligator, lake, etc).
\item Query 2 - {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific autotag and resize to 224x224.
\item Query 3 - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific autotag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
\item Query 4 - {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with 2 specific tags (i.e. alligator AND lake) and resize to 224x224.
\end{itemize}

It is important to note that when querying for images with certain
\textit{autotags}, we also apply a filter using the probability.
For instance, we only retrieve images with a autotag \textit{alligator}
and a probability higher than 92\%.
This probabilities are both present in VDMS (in the form of a property
of the \textit{connection} between the image and that \textit{autotag}),
as well as in MySQL (in the form of a column in the \textit{autotags} table
that links images with tags.

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/concurrency_comparison}
\caption{Concurrency Analysis on Query \textit{1tag\_resize} \- VDMS vs Baseline}
\label{fig:concurrency_comparison}
\end{figure}

% We decided to remove this image because it is redudant with the concurrency
% analysis.
% \begin{figure}[t!]
% \centering
% \includegraphics[width=\columnwidth]{figures/queries_throughput_32}
% \caption{Queries - Throughput - 32 Clients}
% \label{fig:q_throughput_32}
% \end{figure}

% \begin{figure}[t!]
% \centering
% \includegraphics[width=\columnwidth]{figures/queries_throughput_56}
% \caption{Queries - Throughput - 56 Clients}
% \label{fig:q_throughput_56}
% \end{figure}

\begin{figure*}[ht]
\begin{subfigure}{.5\linewidth}
  \centering
  % include first image
  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q1_q2}
  \caption{Query 1 (red), and Query 2 (blue).}
  \label{fig:q1_q2}
\end{subfigure}
\begin{subfigure}{.5\linewidth}
  \centering
  % include second image
  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q3_q4}
  \caption{Query 3 (red), and Query 4 (blue).}
  \label{fig:q3_q4}
\end{subfigure}
\caption{Queries - Throughput - 56 Clients}
\label{fig:q_throughput_56}
\end{figure*}

Image search based on metadata is very expensive on large databases.
Because of the large volume of data, the processing of the retrieve images
is perform in parallel, using multi-core and/or distributed systems.
For instance, a common implementation of an image processing pipeline
would involve the use of distributed processing frameworks
like Hadoop \cite{hadoop} or Spark \cite{spark}.
Consequently, it is key that the data management system used supports
concurrency, providing multiple workers with data in parallel.
The ability to scale with the number of simultaneous clients is key for the
applicability of visual data management systems like VDMS.

>>>>

We have used the YFCC100M dataset to evaluate different aspects of our system.
We use the images in the dataset and its associated metadata to implement
an image-search engine based on properties associated with those images.
This is a very common use-case we have encountered when building
applications such as smart-retail, sports applications, and video summarization,
where the starting point is a large set of data that must be curated
before proceeding with the data processing (such as neural network training).
In order to evaluate the different aspects of the performance on the
image search, we have built a baseline following the methodology
used in the industry, and following what we have done in the past
in order to solve the data search problem,
which we describe in Section \ref{images}.

We also evaluate the performance and scalability of the video management
capabilities offered by VDMS.
Handling video in a general way is a complex task.
Some of this complexity comes from the existence of a variety of open and
proprietary implementations, different encoding techniques and container
formats, and different parameters of the video itself that are
application-dependent, like frames per second, lossy compression, etc).
When it comes to video, ad-hoc solutions have a large number of parameters that
can be tuned.
Together with that, there is no system that enables transactional
operations over videos files in the way VDMS does.
Because of this, we focus our efforts on understanding variations in the
performance of VDMS and its scalability, rather than comparing
it to a baseline that would not represent a fair
comparison for either of the systems.
We describe the results in Section \ref{videos}.

Similarly, we explore the behavior of the feature vector functionality in VDMS,
and an evaluation of the different trade-offs that the systems offers for
application developers. For this, we implemented an image-search application
based on \textit{similarity} search, which we describe in Section \ref{features}.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/comparison_system}
\caption{Comparison Systems: Logical view of the interaction between the client application with VDMS (left) and the baseline system (right).}
\label{fig:systems}
\end{figure*}

@@

The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a large
collection of 100 million public Flickr media objects created to provide free,
sharable multimedia data for research. This dataset contains approximately
99.2 million images and 0.8 million videos with metadata characterized by
25 fields such as the unique identifier, userid,
date the media was taken/uploaded, location in longitude/latitude coordinates,
device type the media was captured, URL to download the media object,
and the Creative Commons license type information.
The YFCC100M dataset also contains \textit{autotags}
provided as a set of comma-separated concepts such as people, scenery, objects,
and animals from 1,570 trained machine learning classifiers ~\cite{Thomee_2016}.
Together with each \textit{autotags}, there is a
probability associated with each tag to indicate certainty of the classification.
This is, an image can have the \textit{autotags} "people", "person", "party",
"outdoor", and each \textit{autotag} assigned will be accompanied by a
probability of that \textit{autotags} being present in that image/frame.
We have also used feature vectors generated for every image and first frame
of every video \cite{features} to implement a similarity search.
Given that there is no standard benchmark oriented towards visual data queries,
we have built a series of queries to filter this dataset that is modeled after
our internal use cases for many of the mentioned applications we have worked
with.

\subsection{Experimental Setup}
\label{setup}

Given that there are no other open-source systems
that implement similar functionality and interfaces as VDMS, as a baseline,
we have implemented an equivalent visual data management system comprised of a
combination of widely available, off-the-shelf components:
MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
Note that this implementation only partially replicates the functionalities
that VDMS offers when it comes to image and metadata handling, built for the
purpose of an ad-hoc image search implementation.
This implementation is based upon internal tools used for
that particular task in the past.
We have implemented a set of client-side applications that take care
of retrieving the components from the different systems, and applies
pre-processing operations when needed.
For visual data workloads, building an approach like the one
implemented as a baseline for this work
is the common practice in the industry \cite{haystack, tao}.

For all our experiments, we use two servers, one hosting a VDMS server and
another hosting the baseline implementation.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with hyper-threading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
has 64GB of DDR4 DRAM.
We decided to run VDMS server in the machine with less DRAM to make
sure MySQL had no disadvantage, and because previous evaluation
indicated smaller footprint in the case of VDMS when
compared to similar baselines based on MySQL.
Other than the difference in DRAM space, machines are identical.
Both servers run Ubuntu 16.04.
The client application running the queries and measuring round-trip time
is connected to the server through a 1GB wired link through
a 10GB back-plane switch, same as both servers.

Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (retrieves metadata and
images) with VDMS (left) and the baseline (right).
The client application was implemented using Python 3 for both VDMS and
the baseline.

It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both 
Apache WebServer and VDMS have direct access. 
In the case of the baseline, metadata is
stored in MySQL using an attached SSD disk.
Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of
comparison with respect to MySQL, which was not designed for
Persistent Memory type of storage.
The benefits of Persistent Memory on metadata operations is left
for another paper, and outside the scope of this evaluation.
For this experiment, in the case of VDMS we simply use a similar
attached SSD disk to store metadata.
Even if PMGD, the graph database used by VDMS, is designed for persistent memory,
it can deliver good performance when using SSDs directly, while still
providing ACID-compliant  transactions.

For the metadata, we built VDMS and MySQL databases
using the YFCC100M dataset with incremental database sizes.
For simplicity, we named the database based on the approximate number of images
it contains, as follows: 1M, 5M, 10M, 50M, 100M.
The VDMS and MySQL databases have comparable number of elements. 
The exact number of images/elements in each database are shown in
Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}.
The differences can be attributed to failures in data 
preparation/loading because of incomplete/inconsistent formatting, 
which is common in large datasets \cite{failures}.
In our set up, that difference is very small: 
less than 0.1\% in terms of number of elements (images and/or metadata information).

\subsubsection{Data Representation}

\textbf{VDMS:}
For each database size, we created an instance of VDMS using the image/video metadata,
the machine-generated \textit{autotags} associated with
each image/video identifier, and the list of 1,570 \textit{autotags}.
Internally, that information is represented as a property graph,
where we have one node for each image, one node for each tag
(always 1,570 tags), and connections between images.
For instance, if an image has four \textit{autotags} assigned,
there will be four connections between that image and
the different nodes for those \textit{autotags}.
The probability the \textit{autotag} is present in an image
is expressed as a property in the \textit{connection} between the two nodes.
Figure~\ref{fig:graph_representation} shows an example on two images,
two \textit{autotags}, and the \textit{connections} between
those \textit{autotags} and the images.
Image id 23143252 has two \textit{autotags} assigned:
\textit{Alligator} with probability 0.285, and \textit{Lake} with probability 0.872.
Image id 86756231, on the other hand, has a single \textit{autotags} assigned:
\textit{Alligator} with probability 0.894.
On average there are 8 tags assigned to each image so
there will be around 8 times more connections than images, as shown
in Table~\ref{table:vdmsnodes}.
Also, each image node will contain multiple properties associated
with it (some of which are listed in Section \ref{dataset}). 
% Using the Python Client Module, we insert 1,570 nodes for the \textit{autotags},
% and then insert nodes for the YFCC media objects (images or videos)
% with the associated metadata (as properties of the image/video).
% Then, for each image, we add a \textit{connection} between the image node 
% and the \textit{autotag} node. The probability of that \textit{autotag}
% being present in that image is stored as a property in that \textit{connection}.
The number of nodes (representing images and \textit{autotags}) 
are dependent on the database size and the \textit{connections} are responsible for
90\% of the elements in each database instances,  
as shown in Table~\ref{table:vdmsnodes}.

It is important to note that we create indexes for the image identifier,
\textit{autotags} properties, and longitude/latitude coordinates 
to enable faster retrieval.

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{figures/graph_representation}
\caption{VDMS Data Representation Using a Property Graph:
Example on two images and 2 \textit{autotags} with
their respective probabilities expressed in the \textit{connection}.
Image id 23143252 has two \textit{autotags} assigned:
\textit{Alligator} with probability 0.285, and \textit{Lake} with probability 0.872.
Similarly, Image id 86756231 has a single \textit{autotags} assigned:
\textit{Alligator} with probability 0.894.}
\label{fig:graph_representation}
\end{figure}

\begin{table}[ht]
\caption{VDMS Database - Number of Elements}
\centering
\begin{tabular}{c c c c}
\hline\hline
DB Name & \# Images & \# Connections & \# TagList\\
\hline
% 100k & 100,000     & 848,432      & 1,570\\
% 500k & 500,000     & 4,249,500    & 1,570\\
1M   & 1,000,000   & 8,503,045    & 1,570\\
5M   & 5,000,000   & 42,505,478   & 1,570\\
10M  & 10,000,000  & 85,040,404   & 1,570\\
50M  & 50,000,000  & 425,162,070  & 1,570\\
100M & 99,205,984  & 895,572,430  & 1,570\\
\hline
\end{tabular}
\label{table:vdmsnodes}
\end{table}

\textbf{MySQL/Baseline:}
Each MySQL database is created in a similar manner as VDMS 
but the data is represented as three tables, following the relational model:
1) \textit{images} table, contains one row per image,
and a column for each property
associated with the images (some of which are listed in Section \ref{dataset});
2) \textit{taglist} table: contains one row per autotag element
(always 1,570 rows);
3) \textit{autotags} table: contains one row per autotag
assigned to an image. Each row contains a foreign key to the
image, a foreign key to the tag, and
the probability assigned to that tag belonging to that image.
Given that there are 8 autotags, on average, per image, the \textit{autotags}
table has around 8 times the number of rows present in the
\textit{metadata} table, as can be seen in Table~\ref{table:mysqltables}.
Using a Python client and simple queries, the \textit{taglist}
table is read from the list of tags with an auto-incremented
\textit{tagid} as a primary key, and the metadata table
is read from the YFCC100M metadata using the identifier as a primary key.
The \textit{autotags} table contains the generated autotags and
probabilities for entries of the \textit{images} table.
To generate the table, we split the \textit{autotags} data for each database
by the image identifier and autotag into new files.
The new files are read into the \textit{autotags} table with the image
identifier and \textit{tagid} as foreign keys.

In an attempt to have the best MySQL configuration possible for this use case, 
we explore several parameters to increase the performance of both loading the data, 
as well executing the queries.
In particular, MySQL optimizes threads and transactions out-of-box, 
but it cannot handle the entire YFCC100M dataset without configuring 
specific parameters.
When creating large databases, a data lock may occur to protect the
data from concurrent updates~\cite{mysql_blog}.
To avoid this mechanism, we increased the buffer pool size to
increase the amount of memory allocated to internal data structures.
It is recommended to set the buffer pool size to 60-80\% of the physical
memory size ~\cite{mysql,mysql_blog}.
However, the time to build a database increased. 
We later changed  the buffer pool size to a multiple of the default value, i.e. 16x,
which produced the best results for loading time.

By default, MySQL uses the available operating system threads to  
execute \textit{n} requests in parallel where \textit{n} is 
the number of background read/write I/O threads.
Setting the respective parameters in the MySQL configuration file can limit the
number of concurrent threads and the number of background threads.
When a limit is set on the number of threads, and no threads are available,
requests will go into a FIFO queue until threads are available to execute
the request ~\cite{mysql,mysql_blog}.
We ran a few experiments investigating the effects of setting a limitation on the
number of concurrent and background threads.
We concluded that the default settings perform better for large databases instead of
setting a limit.
Therefore, we let MySQL to automatically handle the concurrency.

In the case of VDMS, we did not attempt to tune any
parameter to avoid unfairness in the comparison against the baseline.
We use the default parameters provided by the implementation.
For both VDMS and the baseline, we created indexes over the 
properties we used for search, such as name of 
\textit{autotag}, and geo-location values.
Building indexes for the right properties and objects
is basic operation that would be present in any real-world deployment,
and measuring performance without them would lead to useless analysis in our
real-world applications and use cases.

\begin{table}[ht]
\caption{MySQL Database - Number of Rows in each Table}
\centering
\begin{tabular}{c c c c}
\hline\hline
 & \multicolumn{3}{c}{Table}\\
\cline{2-4}
DB Name & images & autotags & taglist\\
\hline
% 100k & 100,000    & 848,912     & 1,570\\
% 500k & 498,707    & 4,241,200   & 1,570\\
1M   & 1,000,000  & 8,508,380   & 1,570\\
5M   & 4,987,379  & 42,425,905  & 1,570\\
10M  & 10,000,000 & 85,095,265  & 1,570\\
50M  & 50,000,000 & 425,446,208 & 1,570\\
100M & 99,206,564 & 896,002,496 & 1,570\\
\hline
\end{tabular}
\label{table:mysqltables}
\end{table}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/db_time_size}
\caption{Time to build and size (in GB) of MySQL and VDMS databases.}
\label{fig:db_time_size}
\end{figure}

@@

One of the first things we noticed is the difference in loading times,
where VDMS outperforms MySQL by a large margin.
This analysis includes the metadata, as the images are stored in a shared filesystem. 
Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
faster than MySQL, and how the speedup is sustained as the database size grows.
Key difference in the build times are attributed to the low-level
implementation of how MySQL reads and stores data from the files and the
optimizations (increased InnoDB pool size, etc.) needed
to handle large datasets such as YFCC100M.
On average, it took MySQL around 3.72x longer to build each database than VDMS.

@@

Another important aspect to note is that
VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
This is space used to store information about each node/connection.
The Graph Database internal to VDMS (called PMGD) was designed for performance,
especially in environments where persistent memory is present.
This design decision comes as a trade-off for storage footprint, which is
noticeable in our results.
VDMS required 30-41\% more storage than MySQL for storing the same amount
of metadata.
This may become a factor if storage is a limitation, but it should also be noted
that even if we have a 41\% increase in metadata size,
metadata accounts for less than 2\% of the overall database size.
For example, the largest database (both metadata and images) we built
(100M) has around 230GB of metadata and 12TB of images.
In systems where persistent memory is a scarce resource,
the increased storage foot print of PMGD may represent a challenge.
On the other hand, persistent memory is expected to be available 
in the order of TBs per server, which should fit the 
metadata of intensive use-cases\cite{IntelXPoint15}.

@@

\subsection{Images Search}
\label{images}

In order to evaluate VDMS and the baseline on our use-case queries,
we implemented 4 queries that filter and retrieve a specific set of images.
We chose these queries because they represent typical use-cases where a
cohort of images is to be retrieved and processed from a large corpus of data.
As we mentioned before, we took this approach due to the lack of standard
benchmarks that are oriented towards visual data retrieval.
We use the metadata associated with the images to filter said images.

We use the \textit{autotags} (as they contain information about the content 
of the image), and geo-location information (latitude/longitude) 
of the images for search and filtering.
Note that, even if we use geo-location for our study, any other property 
assigned to the images can be used to refine the search 
in both VDMS and baseline implementations.
On top of that, and for our use cases, we would like to extract more information
about the content of the image through the use of ML,
such as Convolutional Neural Networks~\cite{cnn}.
For this, we resize the images to 224x224, which is the input layer size for
popular variations of neural networks for object detection on images~\cite{resnet}.

To evaluate the access to metadata and images,
we use the following four queries, modeled after our internal use cases:
\begin{itemize}
\item \textit{q1} - {\bf {\em 1tag}}: Find metadata/images with one specific autotag (i.e. alligator, lake, etc).
\item \textit{q2} - {\bf {\em 1tag\_resize}}: Find metadata/images with one specific autotag and resize to 224x224.
\item \textit{q3} - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with one specific autotag, resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
\item \textit{q4} - {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with two specific autotags (i.e. alligator AND lake), resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
\end{itemize}

It is important to note that when querying for images with certain
\textit{autotags}, we also apply a filter using the probability.
For instance, we only retrieve images with an autotag \textit{alligator}
and a probability higher than 92\%.
These probabilities are both present in VDMS (in the form of a property
of the \textit{connection} between the image and that \textit{autotag}),
as well as in MySQL (in the form of a column in the \textit{autotags} table
that links images with tags).
In the case of VDMS, the query involves a graph traversal query that starts
from the \textit{autotag} node and ends in the images node,
following \textit{connections} between the image and that \textit{autotag}).
In the case of the baseline implementation,
the query involves JOIN operations between the 3 tables.
The implementation of this evaluation, as well as all the queries, are available
under the benchmarks branch of the VDMS project
\footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
and some examples of the queries can be found in the appendix of this
paper.

\begin{figure}[ht]
\includegraphics[width=\columnwidth]{figures/concurrency_comparison}
\caption{Concurrency Analysis on \textit{q2} (\textit{1tag\_resize}).
Hardware concurrency (number of physical cores in each system)
is shown with a blue vertical line (hw-concurrency = 56).
Top figure shows aggregated throughput (transactions per second)
when retrieving only metadata associated with the images, as the number of
concurrent clients increase.
Bottom figure shows aggregated throughput (images per second) when retrieving
resized versions of the images, as the number of concurrent clients increases.}
\label{fig:concurrency_comparison}
\end{figure}

Also, note that the size of the result (number of images retrieved)
is linear with the size of the database. This is, if a query returns 100 images
for the 1M database, it will return around 1000 images for the 10M database.
This poses a problem when evaluating performance as the size of the database increase,
and clearly understanding the measurements.
Because of this reason, we control the number of returned images for all the
databases using the probability of the \textit{autotags} 
(higher probabilities returns less images), so that the queries in 
this experiment return a similar number of images for all database sizes.
In other words, as the size of the database increase, we increase the probability
threshold for the queries. We do this for both VDMS and the baseline, of course.
This way, we remove bottleneck introduced by network bandwidth that would
otherwise over-complicate the understanding of the results.
